SunPy - Python for Solar Physics

### 2014-01-27 1-Introduction.tex

\section{Introduction}
Modern solar physics and astrophysics, along with most of modern science is 
increasingly becoming data-driven. Where observational and numerical data is 
being produced at a faster rate than new hypotheses. For example, NASA's Solar 
Dynamics Observatory (SDO) satellite records over 1 TB of data per day all of 
which is telemetered to the ground and available for analysis.
Solar physics therefore requires increasingly complex software tools to enable 
scientific discovery by analysis of this data. To make meaningful and 
reproducible advances in our understanding of the Sun it is important that the 
software tools to be standardised, easy to use, and transparent, so that the 
community can build upon a common foundation.

SunPy aims to provide a free, open source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breath of high-quality scientific focused packages 
available for it. The Python programming language is a general purpose, 
powerful and easy to learn high-level programming language. Python is one of 
the top ten most popular programming languages in the world according to the 
2014 TIOBE Index 
\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
 and is widely used outside of scientific fields in web development, education 
and even 3D game design.

The development of a package such as SunPy in Python is made possible by the 
rich scientific Python ecosystem of packages available. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language, 
\textit{i.e.} array manipulation, core numerical algorithms and visualisation. 
Ontop of this packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide a tremendous amount of more domain-specific 
functionality.

@@

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project as well as future plans. The latest release of SunPy is available 
in PyPI and can be installed in the usual manner for a Python package
\footnote{http://docs.sunpy.org/en/latest/sunpy/guide/installation/index.html}.

>>>>

\section{Introduction}
Science, including solar physics and astrophysics, is driven by the analysis 
and summarizing of data. Modern advances in sensor technology combined with 
the availability of cheap storage and large bandwidth communications has led to 
a rapid and overwhelming increase in amount of data which scientists are responsible
for analyzing. For example, NASA's Solar 
Dynamics Observatory (SDO) satellite records over 1 TB of data per day. To analyze this 
mountain of data and enable scientific discovery requires increasingly complex software 
tools. Standardised, easy to use,
and transparent software tools are a key enabling technology in the scientific endeavor
so that the community can build upon a common foundation.
%schriste - the opening paragraph still needs work

SunPy aims to provide a free, open source and openly developed software package 
for the analysis and visualisation of solar data. SunPy makes use of the Python 
programming language and the breath of high-quality scientific focused packages 
available for it. The Python programming language is a general purpose, 
powerful and easy to learn high-level programming language. Python is one of 
the top ten most popular programming languages in the world according to the 
2014 TIOBE Index\footnote{\url{http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html}},
and is widely used outside of scientific fields in web development, education 
and `big data' analytics.

The development of a package such as SunPy in Python is made possible by the 
rich scientific Python ecosystem of packages available. Core packages in this 
ecosystem such as \texttt{NumPy}, \texttt{SciPy} and \texttt{matplotlib} 
provide the basic functionality expected of a scientific programming language, 
\textit{i.e.} array manipulation, core numerical algorithms and visualisation. 
Build upon these foundations, packages such as \texttt{Astropy}, \texttt{pandas} and 
\texttt{scikit-image} provide more domain-specific functionality.

@@

The purpose of this paper is to provide an overview of SunPy's current 
capabilities, an overview of the development model and community aspects of the 
SunPy project as well as future plans. The latest release of SunPy (0.4) from
the sunpy website (\url{http://sunpy.org}) or can be installed using 
the Python package index (\mbox{\url{http://pypi.python.org/pypi}}).
%schriste - the mbox prevents latex from breaking the url across different lines.

### 2014-01-27 2-2-Lightcurve.tex

A \texttt{LightCurve} object may be created using a range of different methods. 
Firstly, a \texttt{LightCurve} may be created for a specific instrument based 
on an input time range. In Figure~\ref{code:goes_lc_code}, 
the LightCurve constructor searches a remote site for the GOES X-ray data 
specified by the time interval, downloads the required files and subsequently 
creates the object.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes=lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                       '2011-06-07 10:00')
>>> goes.peek()
>>> trange=TimeRange('2011-06-07 06:00','2011-06-07 07:00')
>>> goes_zoom = goes.truncate(trange)
% the above does not work!
>>> resampled_goes = goes.data.resample("1H")
\end{minted}
\includegraphics[width=14cm]{goes_lightcurve.pdf}
\caption{Example usage of a GOES lightcurve for the time interval 06:00 - 08:00 UT on 
2011 June 7 using a time range, and the result of the \texttt{peek()} method.
A sub-interval is then extracted from the Lightcurve and finally the data is resampled
to 1 hour cadence.}
\label{code:goes_lc}
\end{Figure}

Alternatively, if the data file already exists on the local system, the 
\texttt{Lightcurve} object may instead be initialised using that file as input.
Once the \texttt{Lightcurve} has been created it may be manipulated in 
a variety of ways in order to perform time series analysis. 
Continuing from Figure~\ref{code:lyra_lc}, we show a few simple 
examples of using a LYRA lightcurve.

>>>>

A \texttt{LightCurve} object may be created using a range of different methods. 
Firstly, a \texttt{LightCurve} may be created for a specific instrument based 
on an input time range. In Listing~\ref{code:goes_lc}, 
the LightCurve constructor searches a remote site for the GOES X-ray data 
specified by the time interval, downloads the required files and subsequently 
creates the object.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes=lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                       '2011-06-07 10:00')
>>> goes.peek()
>>> trange=TimeRange('2011-06-07 06:00','2011-06-07 07:00')
%>>> goes_zoom = goes.truncate(trange)
% the above does not work!
>>> resampled_goes = goes.data.resample("1H")
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example usage of a GOES lightcurve for the time interval 06:00 - 08:00 UT on 
2011 June 7 using a time range, and the result of the \texttt{peek()} method.
A sub-interval is then extracted from the Lightcurve and finally the data is resampled
to 1 hour cadence.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

Alternatively, if the data file already exists on the local system, the 
\texttt{Lightcurve} object may instead be initialised using that file as input.
Once the \texttt{Lightcurve} has been created it may be manipulated in 
a variety of ways in order to perform time series analysis.

### 2014-01-27 2-3-Spectra.tex

\subsection{Spectra}\label{sec:spectra}

@@

SunPy aims to provide support to most of solar spectroscopy instrument, however the 
complexity of this kind of datasets makes very challenging to add support to them all 
in a general way.  
Nevertheless, the \texttt{Spectra} object has been designed to be consistent with the 
other data types available in Sunpy.
It was implemented as a wrapper around a data array (numpy ndarray)
because its 1D nature (intensity versus frequency).
Using the \texttt{Spectra} object, Sunpy also offers a 2D array object called \texttt{Spectrogram},
which it is built as a group of spectrum meassured over time.
These two objects were implemented by funding provided by the Astrophysics Research 
Group at Trinity College Dublin.
As yet, Sunpy \texttt{Spectrogram} object supports radio spectra from the e-Callisto 
network\footnote{e-Callisto is an international network of solar radio spectrometers, 
                 all the data is available at: \url{http://www.e-callisto.org/} } 
and STEREO/SWAVES spectrograms.
Like for the \texttt{Lightcurve} object, the \texttt{Spectrogram} object has been
built so each instrument initialises from a sub-class containing the instrument-specific 
functionalities while offering common functionalities to them all to help its analysis, 
this includes:
read the data; %(fits format for e-Callisto, ASCII for STEREO/SWAVES); 
join different time ranges and frequencies; 
frequency dependent background subtraction;  
linearise of frequency axis and automated downsample for proper visualisation on a normal computer screen;  
and point-and-click functionality to save pixels of interest for further analysis.
In addition, this object is equipped with a download data interface for e-Callisto that 
only requires the observatory name and the desired time-range. 
STEREO/SWAVES data can be obtained via the virtual solar observatory (see \S\ref{sec:vso}).

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram} object retrieves
from the online data archive the files that are between the time range specified and
the observatory of interest.  In this case the data is requested from \textit{BIR} which is the code name of the
Rosse Observatory at Birr Castle (Ireland)\footnote{\url{http://www.rosseobservatory.ie}}.
When the data is requested using the \texttt{.from\_range} function, the object merges
into a single spectrogram all the files downloaded, each with 15 minutes observation data 
and in some cases with different files for different frequency ranges.  
In the example shown here BIR observed in two frequency ranges: 20 - 90\,MHz and 55 - 355\,MHz.
Since the data is not evenly spaced in the frequency range, the \texttt{Spectrogram} object
linearise the frequency axis for a better analisys, as shown in the resultant figures.
It is also shown in the example the effect of the implemented background substraction and
\texttt{.peek} method.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
from sunpy.spectra.sources.callisto import CallistoSpectrogram
tstart = "2011-06-07T06:00:00"
tend = "2011-06-07T07:45:00"
callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
callisto_nobg = callisto.substract_bg()
callisto.peek()
callisto_nobg.peek(vmin = 0)
\end{minted}
\includegraphics[width=0.8\columnwidth]{callisto} %\hfill
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\caption{Demonstration of how \texttt{CallistoSpectrogram} object retrieve the data
for the time range and observatory requested, merge it all and removes the background
signal.}
\label{code:spectra}
\end{listing}

>>>>

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work

@@

SunPy aims to provide support to most of solar spectroscopy instrument, however the 
complexity of this kind of datasets makes very challenging to add support to them all 
in a general way.  
Nevertheless, the \texttt{Spectra} object has been designed to be consistent with the 
other data types available in SunPy.
It was implemented as a wrapper around a data array (numpy ndarray)
because its 1D nature (intensity versus frequency).
Using the \texttt{Spectra} object, SunPy also offers a 2D array object called \texttt{Spectrogram},
which it is built as a group of spectrum meassured over time.
These two objects were implemented by funding provided by the Astrophysics Research 
Group at Trinity College Dublin.
As yet, SunPy \texttt{Spectrogram} object supports radio spectra from the e-Callisto 
solar radio spectrometer network(\url{http://www.e-callisto.org/}).
and STEREO/SWAVES spectrograms.
Like for the \texttt{Lightcurve} object, the \texttt{Spectrogram} object has been
built so each instrument initialises from a sub-class containing the instrument-specific 
functionalities while offering common functionalities to them all to help its analysis, 
this includes:
read the data; %(fits format for e-Callisto, ASCII for STEREO/SWAVES); 
join different time ranges and frequencies; 
frequency dependent background subtraction;  
linearise of frequency axis and automated downsample for proper visualisation on a normal computer screen;  
and point-and-click functionality to save pixels of interest for further analysis.
In addition, this object is equipped with a download data interface for e-Callisto that 
only requires the observatory name and the desired time-range. 
STEREO/SWAVES data can be obtained via the virtual solar observatory (see \S\ref{ssec:vso}).

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram} object retrieves
from the online data archive the files that are between the time range specified and
the observatory of interest.  In this case the data is requested from \textit{BIR} which is the code name of the
Rosse Observatory at Birr Castle (Ireland)(\url{http://www.rosseobservatory.ie}).
When the data is requested using the \texttt{.from\_range} function, the object merges
into a single spectrogram all the files downloaded, each with 15 minutes observation data 
and in some cases with different files for different frequency ranges.  
In the example shown here BIR observed in two frequency ranges: 20 - 90\,MHz and 55 - 355\,MHz.
Since the data is not evenly spaced in the frequency range, the \texttt{Spectrogram} object
linearise the frequency axis for a better analisys, as shown in the resultant figures.
It is also shown in the example the effect of the implemented background substraction and
\texttt{.peek} method.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
from SunPy.spectra.sources.callisto import CallistoSpectrogram
tstart = "2011-06-07T06:00:00"
tend = "2011-06-07T07:45:00"
callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
callisto_nobg = callisto.substract_bg()
callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Demonstration of how \texttt{CallistoSpectrogram} object retrieve the data
for the time range and observatory requested, merge it all and removes the background
signal.}
\label{code:spectra}
\end{listing}

### 2014-01-27 3-1-VSO.tex

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
Listing~\ref{code:vso_query_advanced} shows examples of how to make advanced
queries by combining attribute objects.
Two attribute objects can be combined with a logical \texttt{or} operation
using the ``$|$'' operator.
All attribute objects provided to the query as keyword arguments are combined
with a logical \texttt{and} operation.

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> client = vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 10:30'
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend),
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
40
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type
----------           --------             ------  ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA
...
\end{minted}
\caption{Example of querying a single instrument over a time range.}
\label{code:vso_query_simple}
\end{listing}

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

@@

Listing~\ref{code:vso_down} shows how to download the data using the result
from the query.
One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronagraphs, path = pathformat)
\end{minted}
\caption{Example of how to download data using a query result, continuing from
listing~\ref{code:vso_query_advanced}).}
\label{code:vso_down}
\end{listing}

>>>>

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
The also shows how to download the data using the result
from the query. One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

Listing~\ref{code:vso_query_advanced} shows examples of how to make advanced
queries by combining attribute objects.
Two attribute objects can be combined with a logical \texttt{or} operation
using the ``$|$'' operator.
All attribute objects provided to the query as keyword arguments are combined
with a logical \texttt{and} operation.

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from SunPy.net import vso
>>> client = vso.VSOClient()
>>> tstart, tend = '2011/6/7 05:30', '2011/6/7 10:30'
>>> lasco_query = client.query(vso.attrs.Time(tstart, tend),
...                            vso.attrs.Instrument('lasco'))
>>> lasco_query.num_records()
40
>>> lasco_query.show() 
Start time           End time             Source  Instrument  Type
----------           --------             ------  ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO    LASCO       CORONA
2011-06-07 05:43:09  2011-06-07 05:43:29  SOHO    LASCO       CORONA
...
>>> pathformat = '/data/{instrument}/{detector}/{file}.fits'
>>> results = client.get(coronagraphs, path = pathformat)
\end{minted}
\caption{Example of querying a single instrument over a time range and downloading the data}
\label{code:vso_query_simple}
\end{listing}

%schriste - what about just listing the sources as an array to instead of doing a show
%           to make the point that the | query worked appropriately? That would make this
%            listing much shorter.
\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
Querying multiple instruments/missions over the same time range
>>> instrument_condition = ( vso.attrs.Instrument('lasco') |
...                          vso.attrs.Detector('cor1') )
>>> coronagraphs = client.query(vso.attrs.Time(tstart, tend),
...                             instrument_condition)
>>> coronagraphs.num_records()
485
>>> coronagraphs.show()
Start time           End time             Source    Instrument  Type
----------           --------             ------    ----------  ----
2011-06-07 05:35:23  2011-06-07 05:35:48  SOHO      LASCO       CORONA
...
2011-06-07 10:25:43  2011-06-07 10:25:45  STEREO_B  SECCHI      CORONA
2011-06-07 10:30:00  2011-06-07 10:30:01  STEREO_A  SECCHI      CORONA

@@

### 2014-01-27 3-2-HEK.tex

The Sun is an active star and exhibits a wide range of phenomena on many 
different time-scales, length-scales and wavelengths. Observations and metadata 
concerning these phenomena are collected at the Heliophysics Event 
Knowledgebase (HEK).  Entries are generated both by automated algorithms and 
human observers.  Some of the information in the HEK reproduces feature/event 
data elsewhere (for example, the GOES flare catalog), and some is generated by 
the Solar Dynamics Observatory Feature Finding Team \cite{SDOFFT paper???} to
populate the HEK.  The advantage of the HEK to SunPy is that it is a single 
service that provides a homogeneous and well-described interface to a large 
amount of feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was 
developed through support from the European Space Agency Summer of Code in 
Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
Every event type has required and optional properties; the current list of 
events maintained by the HEK and their properties can be found at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

Simple HEK queries consist of start time, an end time, and an event type 
(Listing \ref{code:hek:simple}). Times can be specified in any SunPy-acceptable 
format. Event types are specified as upper case, two letter strings, and are 
identical to the two letter abbreviations found at the HEK website, 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
from sunpy.net import hek
client = hek.HEKClient()
tstart = '2011/08/09 07:23:56'
tend = '2011/08/09 12:40:29'
event_type = 'FL'
result = client.query(hek.attrs.Time(tstart,tend), 
hek.attrs.EventType(event_type))
\end{minted}
\caption{Simple HEK query, returning 19 entries.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods}, or 
FRMs. Example \ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
                      hek.attrs.FRM.Name == 'SSW Latest Events')
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators e.g. AND
and OR. This makes complex queries easy to create: Listing \ref{code:hek:or} 
returns flares west of 50 arcseconds or those that have a peak flux above 
1000.0 (the units of the flux are FRM dependent and are described at the HEK 
website).
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
      (hek.attrs.Event.Coord1 > 50) or (hek.attrs.FL.PeakFlux > 1000.0))
\end{minted}
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}), but the optional attributes 
are FRM dependent.  If a FRM does not have one of the optional attribute, a 
'None' is returned by the \texttt{hek} module.

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity, bringing powerful 
feature/event search capabilities to SunPy. It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena (e.g. flares, 
radio burst) on many 
different time-scales, length-scales and wavelengths. Observations and metadata 
concerning these phenomena are collected at the Heliophysics Event 
Knowledgebase (HEK).  Entries are generated both by automated algorithms and 
human observers.  Some of the information in the HEK reproduces feature/event 
data elsewhere (for example, the GOES flare catalog), and some is generated by 
the Solar Dynamics Observatory Feature Finding Team \cite{SDOFFT paper???} to
populate the HEK.  The advantage of the HEK is it provides a homogeneous and 
well-described interface to a large 
amount of feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was 
developed with support from the European Space Agency Summer of Code in 
Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
The current list of 
events maintained by the HEK and their properties can be found at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

Simple HEK queries consist of start time, an end time, and an event type 
(Listing \ref{code:hek:simple}). Event types are specified as upper case, 
two letter strings, and are 
identical to the two letter abbreviations found at the HEK website, 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import hek
>>> client = hek.HEKClient()
>>> tstart = '2011/08/09 00:00:00'
>>> tend = '2011/08/10 00:00:00'
# FL = flare
>>> event_type = 'FL'
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type))
>>> len(result)
52
\end{minted}
\caption{Example usage of the \texttt{hek} module showing a simple HEK search for solar flares
which occurred on the 2011/08/09.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods} ( 
FRMs). Example \ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
                      hek.attrs.FRM.Name == 'SSW Latest Events')
>>> len(result)
9
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

HEK comparisons can be combined using Python’s logical operators e.g. AND
and OR. This makes complex queries easy to create: Listing \ref{code:hek:or} 
returns flares west of 50 arcseconds or those that have a peak flux above 
1000.0 (the units of the flux are FRM dependent and are described at the HEK 
website).
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type),
...                       (hek.attrs.Event.Coord1 > 50) 
...                       or (hek.attrs.FL.PeakFlux > 1000.0))
\end{minted}
\caption{HEK query using the 'or' operator.}
\label{code:hek:or}
\end{listing}
All FRMs report the required feature attributes (as described at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}), but the optional attributes 
are FRM dependent.  If a FRM does not have one of the optional attribute, a 
'None' is returned by the \texttt{hek} module.

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.
%schriste - this is a weird way to end this section, either show how it is done or 
% do not mention it, i think.

### 2014-01-27 3-3-HELIO.tex
Minor edit

The HELiophysics Integrated Observatory 
(HELIO)\footnote{\url{http://helio-vo.eu}} has put together a list of web 
services which allows scientists to query and discover data throughout 
the heliosphere (\textit{e.g.}, solar, planetary, inter-planetary data, 
\ldots)\cite{dps2012}. 
HELIO has been built with a Service-Oriented Architecture, 
\textit{i.e.} its capabilities are split into a number of tasks that are 
implemented as separate services. 
HELIO comprises nine different public services, which allows scientists
to search in different catalogues of registered events, solar features,
data from instruments in the heliosphere and other information like planets or 
spacecraft position in time. 
Additionally, HELIO provides a service that uses a propagation model to link 
the data in different points of the solar system by its original nature 
(\textit{e.g.}, Earth auroras are a signature of magnetic field disturbances 
produced few days before on the Sun).
In addition to the primary, web-based interface to HELIO
(\url{http://hfe.helio-vo.eu}), the services can be separately accessed.

@@

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helio import hec
>>> hc = hec.Client()

Desired time range and event type
>>> tstart, tend = '2011-06-07T06:00:00', '2011-06-07T12:00:00'
>>> event_type = 'cme'

>>>>

The HELiophysics Integrated Observatory 
(HELIO, \url{http://helio-vo.eu}) has put together a list of web 
services which allows scientists to query and discover data throughout 
the heliosphere (\textit{e.g.}, solar, planetary, inter-planetary data, 
\ldots)\citep{dps2012}.
HELIO has been built with a Service-Oriented Architecture, 
\textit{i.e.} its capabilities are split into a number of tasks that are 
implemented as separate services. 
HELIO comprises nine different public services, which allows scientists
to search in different catalogues of registered events, solar features,
data from instruments in the heliosphere and other information like planets or 
spacecraft position in time. 
Additionally, HELIO provides a service that uses a propagation model to link 
the data in different points of the solar system by its original nature 
(\textit{e.g.}, Earth auroras are a signature of magnetic field disturbances 
produced few days before on the Sun).
In addition to the primary, web-based interface to HELIO
(\url{http://hfe.helio-vo.eu}), the services can be separately accessed.

@@

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helio import hec
>>> hc = hec.Client()
>>> tstart, tend = '2011-06-07T06:00:00', '2011-06-07T12:00:00'
>>> event_type = 'cme'

### 2014-01-27 3-5-Database.tex
typos

\subsection{The File Database}
Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organized and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
Sunpy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of the 2013 
Google Summer of Code.

The database sub-package makes use of a database supported by
SQLAlchemy\footnote{\url{http://www.sqlalchemy.org}}. This library was chosen
since it supports many SQL dialects. If SQLite is choosen, the database is stored as a single file which is
created automatically. This solution makes it possibly to easily backup, copy and distribute
a database file. A server-based database, on the other hand, could be used by a collaboration 
who work together from different machines on the
same data: a central database server stores all data and the clients connect to
it to read or write data.

@@

Using the database package, one uses the \texttt{Database} class to connect
with a database and perform operations on it. Figure~\ref{code:database} shows how to connect
to an in-memory database and download data from the \textsc{VSO}. These entried are
automatically added to the database. The function \texttt{len} is used to get the number of
records. The function \texttt{display\_entries} displays an iterable of database entries
in a nicely-formatted \textsc{ASCII} table. The headlines correspond to the
attributes of the respective database entries.

>>>>

\subsection{The File Database}
Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organized and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
SunPy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of the 2013 
Google Summer of Code.

The database sub-package makes use of a database supported by
SQLAlchemy(\url{http://www.sqlalchemy.org}). This library was chosen
since it supports many SQL dialects. If SQLite is choosen, the database is stored as a single file which is
created automatically. This solution makes it possible to easily backup, copy and distribute
a database file. A server-based database, on the other hand, could be used by a collaboration 
who work together from different machines on the
same data: a central database server stores all data and the clients connect to
it to read or write data.

@@

Using the database package, one uses the \texttt{Database} class to connect
with a database and perform operations on it. Listing~\ref{code:database} shows how to connect
to an in-memory database and download data from the \textsc{VSO}. These entries are
automatically added to the database. The function \texttt{len} is used to get the number of
records. The function \texttt{display\_entries} displays an iterable of database entries
in a nicely-formatted \textsc{ASCII} table. The headlines correspond to the
attributes of the respective database entries.
